---
title: 'Reproducible Research: Peer Assessment 1'
author: 'Author: rrwiren'
date: "Current date ;) `r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    keep_md: yes
  pdf_document: default
  word_document: default
---

---
Check that home directory is default
> path.expand("~/")
[1] "C:/Users/username/Documents/"
---

### Loading libraries and some housekeeping before starting

```{r, warning=FALSE, message=FALSE}
library(ggplot2)
library(Hmisc)
library(DMwR)
setwd("~/GitHub/RepData_PeerAssessment1")
```

## Loading and preprocessing the data

```{r, results='markup', warning=TRUE, message=TRUE}
if(!file.exists('activity.csv')){
        unzip('activity.zipMiss <- function(x){sum(is.na(x))/length(x)*100}p')
}
activity <- read.csv('activity.csv')
```

## What is mean total number of steps taken per day?

#### 1 - Make a histogram of the total number of steps taken each day

```{r} 
totalStepsDay <- tapply(activity$steps, activity$date, sum, na.rm=T)
qplot(totalStepsDay, xlab='Total steps per day', ylab='Count with binwidth = 1000', binwidth=1000)
```

#### 2 - Calculate and report the mean and median total number of steps taken per day

```{r}
totalStepsDayMean <- mean(totalStepsDay)
summary(totalStepsDayMean)
```
```{r}
totalStepsDayMedian <- median(totalStepsDay)
summary(totalStepsDayMedian)
```

*The `mean` is **`r totalStepsDayMean`** and the `median` is **`r totalStepsDayMedian`**.*

## What is the average daily activity pattern?

```{r}
totalAvgSteps <- aggregate(x=list(meanSteps=activity$steps), by=list(interval=activity$interval), FUN=mean, na.rm=T)
```

##### 1 - Time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis) and 
##### 2 - Showing which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r}
totalMostStepsInterval <- which.max(totalAvgSteps$meanSteps)
ggplot(data=totalAvgSteps, aes(x=interval, y=meanSteps)) +
    geom_line(colour = "brown") +
        geom_vline(aes(xintercept=as.numeric(activity$interval[totalMostStepsInterval])), linetype=4, colour="grey") +
        geom_segment(aes(x = 1050, y = 195, xend = 925, yend = 180), arrow = arrow(length = unit(0.5, "cm"))) +
    xlab("5-minute interval") +
    ylab("Average number of steps taken across all days") 
```

```{r}
totalMostStepsInterval <- which.max(totalAvgSteps$meanSteps)
totalAvgSteps[totalMostStepsInterval, ]
totalTimeMostSteps <- gsub("([0-9]{1,2})([0-9]{2})", "\\1:\\2", totalAvgSteps[totalMostStepsInterval,'interval'])
```

*The 5-minute interval "**`r totalTimeMostSteps`**", contains the maximum number of average steps across all the days. Highlighted also by a grey dotted vertical line.*

## Imputing missing values

##### 1 - Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)

```{r}
missingValues <- length(which(is.na(activity$steps)))
print(missingValues)
```

##### 2 - Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.

```{r}
totalStepsDayMean.i <- activity
totalStepsDayMean.i$steps <- impute(activity$steps, fun = mean)
```

##### 3 - Create a new dataset that is equal to the original dataset but with the missing data filled in.

```{r}
activity.i <- activity
activity.i$steps <- impute(activity$steps, fun=mean)
summary(activity.i)
```

##### 4 - Make a histogram of the total number of steps taken each day and report the mean and median total number of steps taken per day. Do these values differ from the estimates from the first part of the assignment? What is the impact of imputing missing data on the estimates of the total daily number of steps?

```{r}
totalStepsDay.i <- tapply(activity.i$steps, activity.i$date, sum)
qplot(totalStepsDay.i, xlab='Total steps per day (imputed)', ylab='Count with binwith = 1000', binwidth=1000)
```

```{r}
totalStepsDayMean.i <- mean(totalStepsDay.i)
totalStepsDayMedian.i <- median(totalStepsDay.i)
```

*The `mean` is **`r totalStepsDayMean.i`** and the `median` is **`r totalStepsDayMedian.i`**.*

```{r}
diff_mean <- totalStepsDayMean.i - totalStepsDayMean
diff_median <- totalStepsDayMedian.i - totalStepsDayMedian
```
*The `mean` difference is **`r diff_mean`** and the `median` difference is **`r diff_median`** between imputed and non-imputed.*

## Are there differences in activity patterns between weekdays and weekends?

##### 1 - Create a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.

```{r}
activity.i$dayType <- ifelse(as.POSIXlt(activity.i$date)$wday %in% c(0,6), 'Weekend', 'Weekday')
``` 

```{r}
meanWeekday <- activity.i$steps[activity.i$dayType == "Weekday"]
summary(meanWeekday)
``` 
```{r}
meanWeekend <- activity.i$steps[activity.i$dayType == "Weekend"]
summary(meanWeekend)
``` 

##### 2 - Panel plot containing a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis).

```{r}
averagedActivity.i <- aggregate(steps ~ interval + dayType, data=activity.i, mean)
ggplot(averagedActivity.i, aes(interval, steps)) + 
    geom_line(colour = "brown") +
        geom_hline(aes(yintercept=50), linetype=4, colour="grey") +
    facet_grid(dayType ~ .) +
    xlab("5-minute interval") + 
    ylab("Avarage number of steps")
```


##### From comparing the two above charts we can see that activity starts earlier in the weekdays and we can also note that there are more overall activity in the weekends. The grey dotted horisontal line highlights 50 steps per interval.
=======
---
title: 'Reproducible Research: Peer Assessment 1'
author: 'Author: rrwiren'
date: "Current date ;) `r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    keep_md: yes
  pdf_document: default
  word_document: default
---

---
Check that home directory is default
> path.expand("~/")
[1] "C:/Users/username/Documents/"
---

### Loading libraries and some housekeeping before starting

```{r, warning=FALSE, message=FALSE}
library(ggplot2)
library(Hmisc)
library(DMwR)
setwd("~/GitHub/RepData_PeerAssessment1")
```

## Loading and preprocessing the data

```{r, results='markup', warning=TRUE, message=TRUE}
if(!file.exists('activity.csv')){
        unzip('activity.zipMiss <- function(x){sum(is.na(x))/length(x)*100}p')
}
activity <- read.csv('activity.csv')
```

## What is mean total number of steps taken per day?

#### 1 - Make a histogram of the total number of steps taken each day

```{r} 
totalStepsDay <- tapply(activity$steps, activity$date, sum, na.rm=T)
qplot(totalStepsDay, xlab='Total steps per day (imputed)', ylab='Count with binwidth = 1000', binwidth=1000)
```

#### 2 - Calculate and report the mean and median total number of steps taken per day

```{r}
totalStepsDayMean <- mean(totalStepsDay)
summary(totalStepsDayMean)
```
```{r}
totalStepsDayMedian <- median(totalStepsDay)
summary(totalStepsDayMedian)
```

*The `mean` is **`r totalStepsDayMean`** and the `median` is **`r totalStepsDayMedian`**.*

## What is the average daily activity pattern?

```{r}
totalAvgSteps <- aggregate(x=list(meanSteps=activity$steps), by=list(interval=activity$interval), FUN=mean, na.rm=T)
```

##### 1 - Time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis) and 
##### 2 - Showing which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r}
totalMostStepsInterval <- which.max(totalAvgSteps$meanSteps)
ggplot(data=totalAvgSteps, aes(x=interval, y=meanSteps)) +
    geom_line(colour = "brown") +
        geom_vline(aes(xintercept=as.numeric(activity$interval[totalMostStepsInterval])), linetype=4, colour="grey") +
        geom_segment(aes(x = 1050, y = 195, xend = 925, yend = 180), arrow = arrow(length = unit(0.5, "cm"))) +
    xlab("5-minute interval") +
    ylab("Average number of steps taken across all days") 
```

```{r}
totalMostStepsInterval <- which.max(totalAvgSteps$meanSteps)
totalAvgSteps[totalMostStepsInterval, ]
totalTimeMostSteps <- gsub("([0-9]{1,2})([0-9]{2})", "\\1:\\2", totalAvgSteps[totalMostStepsInterval,'interval'])
```

*The 5-minute interval "**`r totalTimeMostSteps`**", contains the maximum number of average steps across all the days. Highlighted also by a grey dotted vertical line.*

## Imputing missing values

##### 1 - Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)

```{r}
missingValues <- length(which(is.na(activity$steps)))
print(missingValues)
```

##### 2 - Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.

```{r}
totalStepsDayMean.i <- activity
totalStepsDayMean.i$steps <- impute(activity$steps, fun = mean)
```

##### 3 - Create a new dataset that is equal to the original dataset but with the missing data filled in.

```{r}
activity.i <- activity
activity.i$steps <- impute(activity$steps, fun=mean)
summary(activity.i)
```

##### 4 - Make a histogram of the total number of steps taken each day and report the mean and median total number of steps taken per day. Do these values differ from the estimates from the first part of the assignment? What is the impact of imputing missing data on the estimates of the total daily number of steps?

```{r}
totalStepsDay.i <- tapply(activity.i$steps, activity.i$date, sum)
qplot(totalStepsDay.i, xlab='Total steps per day (imputed)', ylab='Count with binwith = 1000', binwidth=1000)
```

```{r}
totalStepsDayMean.i <- mean(totalStepsDay.i)
totalStepsDayMedian.i <- median(totalStepsDay.i)
```

*The `mean` is **`r totalStepsDayMean.i`** and the `median` is **`r totalStepsDayMedian.i`**.*

```{r}
diff_mean <- totalStepsDayMean.i - totalStepsDayMean
diff_median <- totalStepsDayMedian.i - totalStepsDayMedian
```
*The `mean` difference is **`r diff_mean`** and the `median` difference is **`r diff_median`** between imputed and non-imputed.*

## Are there differences in activity patterns between weekdays and weekends?

##### 1 - Create a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.

```{r}
activity.i$dayType <- ifelse(as.POSIXlt(activity.i$date)$wday %in% c(0,6), 'Weekend', 'Weekday')
``` 

```{r}
meanWeekday <- activity.i$steps[activity.i$dayType == "Weekday"]
summary(meanWeekday)
``` 
```{r}
meanWeekend <- activity.i$steps[activity.i$dayType == "Weekend"]
summary(meanWeekend)
``` 

##### 2 - Panel plot containing a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis).

```{r}
averagedActivity.i <- aggregate(steps ~ interval + dayType, data=activity.i, mean)
ggplot(averagedActivity.i, aes(interval, steps)) + 
    geom_line(colour = "brown") +
        geom_hline(aes(yintercept=50), linetype=4, colour="grey") +
    facet_grid(dayType ~ .) +
    xlab("5-minute interval") + 
    ylab("Avarage number of steps")
```

##### From comparing the two above charts we can see that activity starts earlier in the weekdays and we can also note that there are more overall activity in the weekends. The grey dotted horisontal line highlights 50 steps per interval.
